---
title: "Change the title to one of your choice"
author: "Toan Pham"
output:
  html_document: default
  pdf_document: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE, fig.height = 4, fig.width = 6 )
```


```{r}
library(tidyverse)
library(magrittr) # needs to be run every time you start R and want to use %>%
library(dplyr) 
library(rlist)
```


```{r, CACHE = TRUE}
# read in data 
Colleges1 <- read.csv("https://ed-public-download.app.cloud.gov/downloads/Most-Recent-Cohorts-Scorecard-Elements.csv")
Colleges2 <- read.csv("https://ed-public-download.app.cloud.gov/downloads/Most-Recent-Cohorts-Treasury-Elements.csv")
Colleges3 <- read.csv("https://ed-public-download.app.cloud.gov/downloads/Most-Recent-Cohorts-All-Data-Elements.csv")
# dimensions should be:
# Colleges1 7112 by 190
# Colleges2 7112 by 92
# Colleges3 7112 by 1978
#check that this is the case
```

Type introductory paragraph. 
 
                                       
```{r}
#First figure (table or graph)      
Project2 <- Colleges3 %>% 
  select(STABBR,INSTNM,SAT_AVG, contains("PCIP"),contains("DEBT"),MN_EARN_WNE_P10,MD_EARN_WNE_P10,  contains("MN_WNE_INC"),COUNT_WNE_P6,MD_EARN_WNE_P6,TUITIONFEE_IN,MN_EARN_WNE_MALE0_P10,MN_EARN_WNE_MALE1_P10,MN_EARN_WNE_MALE0_P6,MN_EARN_WNE_MALE1_P6)
TuitionData1 <- Project2 %>% 
  select(STABBR,TUITIONFEE_IN,MD_EARN_WNE_P6) %>%
  mutate_at(c("TUITIONFEE_IN","MD_EARN_WNE_P6") , as.character) %>%
  mutate_at(c("TUITIONFEE_IN","MD_EARN_WNE_P6") , as.numeric)
ggplot(data = TuitionData1) +
  geom_point(mapping = aes(x = log2(TUITIONFEE_IN), y = MD_EARN_WNE_P6),shape = 21,color = "black", size = 1, stroke = 1, fill = "white") +
  geom_smooth(mapping = aes(x = log2(TUITIONFEE_IN), y = MD_EARN_WNE_P6), size = 2, color = "blue") +
  theme_bw() +
  xlab("Tuition fee") +
  ylab("Earnings after 6 years")
```


one or two paragraph discussion of first figure
There is a clear trend here: as tuition fee increases, earnings after 6 years also increases. This mean that tuition fee directly correlate with earnings after 6 years.
As we observe more closely, we can also see that the rate at which the earnings increase is not constant. To be more specific, while it starts off slowly, the slope get higher and  higher as tuition fee increase. This means that below a certain tuition fee cap, the change in earnings does not really matter. However, at more prestigious universities, the difference in earnings is more notable.
For anyone interested,before I used log2 to make the graph clearer, I did calculate the tuition fee breakpoint, the breakpoint is at around $10000/years.


```{r}
#Second figure (figure or graph)
TuitionData2 <- Project2 %>% 
  select(STABBR,TUITIONFEE_IN,MD_EARN_WNE_P6,MD_EARN_WNE_P10) %>%
  mutate_at(c("TUITIONFEE_IN","MD_EARN_WNE_P6","MD_EARN_WNE_P10") , as.character) %>%
  mutate_at(c("TUITIONFEE_IN","MD_EARN_WNE_P6","MD_EARN_WNE_P10") , as.numeric) %>%
  group_by(STABBR) %>%
  summarize(medianFee = median(TUITIONFEE_IN,na.rm = TRUE),
            medianEarnafter6 = median(MD_EARN_WNE_P6, na.rm = TRUE),
            medianEarnafter10 = median(MD_EARN_WNE_P10, na.rm = TRUE)) %>%
  pivot_longer(cols = c(medianEarnafter6,medianEarnafter10), 
                                      names_to = "year", values_to="earnings") %>%
  filter(STABBR %in% c("CT","DC","IA", "MA","NE","NH","PA","RI","TN","VT"))
ggplot(data = TuitionData2) +
  geom_bar(mapping = aes(x =STABBR , y = earnings), stat = "identity",show.legend = FALSE) +
  geom_hline((aes(yintercept = mean(earnings))), color = "red", size = 1) +
  facet_wrap(~ year) +
  coord_cartesian(ylim=c(25000,48000)) +
  theme_bw() +
  ylab("Average earnings after X years") +
  xlab("States") +
  labs(title = "Top 10 states with highest earnings after 6 and 10 years")

```

one or two paragraph discussion of second figure
Right from the start, we can clearly see the difference between earnings after 6 years and 10 years. The state with the biggest growth is DC and the one with the slowest growth seems to be Tennessee. Overall, the average earnings is around $34000.
In conclusion, for students who want to make good money right away, Massachusettss is a great place to start. For students who prefer better results in the long run, Washington DC is the best choice. In general, these are the 10 states with highest earnings so they all are solid choices.


```{r}
#Third figure (table or graph)
GenderData <- Project2 %>% 
  select(MN_EARN_WNE_MALE0_P10,MN_EARN_WNE_MALE1_P10,MN_EARN_WNE_MALE0_P6,MN_EARN_WNE_MALE1_P6) %>%
  mutate_at(c("MN_EARN_WNE_MALE0_P10","MN_EARN_WNE_MALE1_P10","MN_EARN_WNE_MALE0_P6","MN_EARN_WNE_MALE1_P6") , as.character) %>%
  mutate_at(c("MN_EARN_WNE_MALE0_P10","MN_EARN_WNE_MALE1_P10","MN_EARN_WNE_MALE0_P6","MN_EARN_WNE_MALE1_P6") , as.numeric) %>%
  rename(
    female = MN_EARN_WNE_MALE0_P10,
    male = MN_EARN_WNE_MALE1_P10
    ) %>%
  pivot_longer(cols = c(female,male), 
                                      names_to = "gender", values_to="earnings") %>%
  group_by(gender) %>%
  summarize(medianEarnings = median(earnings,na.rm = TRUE)); GenderData

  
```

one or two paragraph discussion of third figure
Not much to say here, male will earns more than female in the long run. The difference in median is around $10000



```{r}
#Fourth figure (table or graph) - optional
DebtData <- Project2 %>% 
  select(GRAD_DEBT_MDN,MD_EARN_WNE_P6,LO_INC_DEBT_MDN,MD_INC_DEBT_MDN,HI_INC_DEBT_MDN) %>%
  mutate_at(c("GRAD_DEBT_MDN","MD_EARN_WNE_P6","LO_INC_DEBT_MDN","MD_INC_DEBT_MDN","HI_INC_DEBT_MDN") , as.character) %>%
  mutate_at(c("GRAD_DEBT_MDN","MD_EARN_WNE_P6","LO_INC_DEBT_MDN","MD_INC_DEBT_MDN","HI_INC_DEBT_MDN") , as.numeric) %>%
  rename(
    Low_Income = LO_INC_DEBT_MDN,
    Middle_Income = MD_INC_DEBT_MDN,
    High_Income = HI_INC_DEBT_MDN
    ) %>%
  pivot_longer(cols = c(Low_Income,Middle_Income,High_Income), 
                                      names_to = "Income_Level", values_to="Median_Debt")
ggplot(data = DebtData) +
  geom_point(mapping = aes(x = log2(Median_Debt), y = MD_EARN_WNE_P6, fill = Income_Level),shape = 21,color = "black", size = 1, stroke = 1) +
  geom_smooth(mapping = aes(x = log2(Median_Debt), y = MD_EARN_WNE_P6,color = Income_Level), size = 1) +
  theme_bw() + xlab("log of Median debt") + ylab("median earning after 6 years")
  

```

one or two paragraph discussion of fourth figure
                     
In generally, it is worth to have higher debt if the students can get into more prestige schools. These schools can help students to make more connections and get more job opportunities. However, there is a cut in which it is no longer true. The cut happens when log of median debt equals 13 and when log debt is larger than 15. This can be explained that when the debt is too much, it is hard to focus on schools thus 



```{r}
#Fifth figure (table or graph) - optional
Colleges_SAT <- Colleges1 %>% select(INSTNM, STABBR, SAT_AVG)


Colleges_SAT$SAT_AVG <- as.character(Colleges_SAT$SAT_AVG)
Colleges_SAT$SAT_AVG <- as.numeric(Colleges_SAT$SAT_AVG)

Colleges_money <- Colleges3 %>% select(INSTNM, MD_EARN_WNE_P6, DEBT_MDN_SUPP)

Colleges_SAT_complete <- full_join(Colleges_SAT,Colleges_money)

Colleges_SAT_complete$MD_EARN_WNE_P6 <- as.character(Colleges_SAT_complete$MD_EARN_WNE_P6)

Colleges_SAT_complete$MD_EARN_WNE_P6 <- as.numeric(Colleges_SAT_complete$MD_EARN_WNE_P6)

highest_sat <- Colleges_SAT_complete %>% group_by(STABBR) %>% summarize(mean_sat = mean(SAT_AVG, na.rm = TRUE)) %>% arrange(desc(mean_sat))

head(Colleges_SAT_complete)
#According to highest_sat, top 10 states with highest SAT average are DC, RI, NH, UT, MA, WY, CA, MN, NY, WA

ggplot(Colleges_SAT_complete, aes(x = SAT_AVG,y = MD_EARN_WNE_P6, stroke = 1, alpha = 0.1)) + geom_point() + xlab("SAT average") + ylab("Median earnings of students working") + theme_bw() + geom_smooth(method = "lm") + theme(legend.position = "none")
```

one or two paragraph discussion of fifth figure
The graph above shows the correlation between SAT average and median earnings of students working 6 years after graduation. The linear model establishes a positive trend, indicating that with higher SAT score, students tend to make more money after graduation. This makes sense as with high SAT score, students have more chance to get into better schools thus make more connections and be more prepared when it comes to getting jobs. However, there are a few outliers. There are students who make above $60.000 with a 1100 SAT average. Numerous factors could be responsible for this: some students may do bad in one test but excel at college, some students may have talents in different fields and utilize it, some might have a wealthy family, etc. All in all, those are just a small number of outliers and will not change the fact that SAT score plays a role in students' financial success.

## Appendix 

```{r}
#Additional figure(s) if desired
```